---
title: "Stable Bands: Basic Implementation"
author: Jonas Elmerraji (jonaselm@hey.com)
output: html_notebook
---

This notebook contains a full, clean implementation of stable bands, including weekly projection to daily and rolling estimation. This 
software is provided as-is for informational purposes with no implied license. 

As implemented, this code requires STABLE for R, available from http://robustanalysis.com/. A free open source alternative would be
to use the `alphastable` package in place of the stable.fit.mle.restricted() function for parameter estimation, and then using the `stabledist` 
package in place of the qstable() function for quantile calculation. 

```{r}
library(tidyverse)
library(tidyquant)
library(stable)
library(roll)
```

Load price data from your provider into your R environment as a tibble called price_data. The code below assumes price_data has the following standard Bloomberg column names: date, PX_LAST, PX_OPEN, PX_HIGH, PX_LOW.

```{r}
#S&P 500 
#load("data/SPX.Rdata")

```

Add stable bands to your price_data object at each observation.

```{r}
# Calculate weekly invariants
px_weekly <- price_data |> tq_transmute(select = PX_LAST:PX_LOW,
                                        mutate_fun = to.period,
                                        period = "weeks")

# Calculate log-returns
price_data <- price_data |> select(date, PX_LAST) |> mutate(log_return = log(PX_LAST/lag(PX_LAST)))
px_weekly <- px_weekly |> select(date, PX_LAST) |> mutate(log_return = log(PX_LAST/lag(PX_LAST)))

# Calculate 20-period SMA
px_weekly <- px_weekly |> mutate(SMA = SMA(PX_LAST, n=20))
price_data <- price_data |> mutate(SMA = SMA(PX_LAST, n=20))

# Calculate 20-period sd
px_weekly <- px_weekly |> mutate(sd = roll_sd(PX_LAST, 20),
                                 sd_logret = roll_sd(log_return, 20))

price_data <- price_data |> mutate(sd = roll_sd(PX_LAST, 20),
                                 sd_logret = roll_sd(log_return, 20))


# Initialize parameter placeholder
px_weekly <- px_weekly |> mutate(
  alpha = NA,
  beta = NA, 
  gamma = NA, 
  delta = NA
)

# Estimate rolling stable parameters
# Loop through data, taking full filtration
# Here, we fix beta, delta and only estimate alpha, gamma parameter

for (i in 30:nrow(px_weekly)){ # Starting at 3: 2 because we're estimating 2 parameters to start, plus 1 to ensure no lookahead (only fit on prior data)
  chunk <- px_weekly[2:(i-1), ]$log_return #omit leading NA
  param_estimates <- stable.fit.mle.restricted(chunk, theta = c(0,0,0,0), restriction = c(0,1,0,1), param = 1) # Note, parameterization doesn't matter since beta = 0
  
  # Store results
  px_weekly[i,]$alpha <- param_estimates[1]
  px_weekly[i,]$beta <- param_estimates[2]
  px_weekly[i,]$gamma <- param_estimates[3]
  px_weekly[i,]$delta <- param_estimates[4]
  
}

# Quantile of interest - percent of excursions
qtile <- 0.05
norm_qtile <- qnorm(qtile, lower.tail = F)
qtile <- qtile/2

### Project weekly to daily
# Columns to keep from px_weekly
keeps <- px_weekly %>% dplyr::select(date, alpha, gamma)

# Merge and locf observations each week to fill missing days
price_data <- price_data %>% left_join(keeps, by = "date") %>% na.locf()

# Set proj_factor = hs; 
proj_factor <- 5*(1/5)

# Project results
price_data <- price_data |> na.omit() |> rowwise() |> 
  mutate(upside_band_log = qstable(1-qtile, alpha, beta = 0, (proj_factor*(gamma^alpha))^(1/alpha), 0, param = 1),
                                   downside_band_log = qstable(qtile, alpha, beta = 0, (proj_factor*(gamma^alpha))^(1/alpha), 0, param = 1),
         upside_band_ret = exp(upside_band_log)-1, downside_band_ret = exp(downside_band_log)-1,
         upside_band = SMA*(1+upside_band_ret), downside_band = SMA*(1+downside_band_ret))

```

